\section{Applying the Fundamental Theorem of Calculus to a Limit Problem}
\label{sec:ftc_problem}			

\subsection*{Recommended Tutorials:}
\begin{itemize}
	\item Tutorial \ref{chp:limits}, pg. \pageref{chp:limits}
	\item Tutorial \ref{chp:definite_and_indefinite_Integrals}, pg. \pageref{chp:definite_and_indefinite_Integrals}
\end{itemize}

\subsection*{Introduction:}

In this activity, students will be asked to demonstrate their knowledge of the Fundamental Theorem of Calculus in order to evaluate \[\displaystyle\lim_{x \rightarrow 0} \dfrac{\displaystyle\int_{0}^x \sin(t^2) \; dt}{x^3}.\]

\subsection*{Exercises:}

\begin{enumerate}
    \item   Plot the graph of $\dfrac{\displaystyle\int_{0}^x \sin(t^2) \; dt}{x^3}$.  Graphically determine the value of \[\displaystyle\lim_{x \rightarrow 0} \dfrac{\displaystyle\int_{0}^x \sin(t^2) \; dt}{x^3}.\]
    \item   Can we evaluate $\displaystyle\int_{0}^x \sin(t^2) \; dt$?  If not, why?  If so, what is it? 
    \marginnote[-4cm]{It is important to be able to use theorems from both Calculus I and II in order to solve problems.  Maple is a great tool to help you find an answer to a problem; however, Maple graphs do not constitute a proof.  Even when you know the answer, it is important to learn to verify your Maple result.}
    \item   What rule do we need to use in order to evaluate the limit?  State the theorem, including conditions under which we can apply the theorem.
    \item   Apply the theorem.  What happens now?
    \marginnote[-2.5cm]{For questions that require an explanation, use a new paragraph in Maple.}
     \marginnote{Always be careful when you apply a theorem.  Ensure that the hypotheses are fully satisfied before you apply the theorem.}
    \item   What is the value of the limit?
    \item   Write your answer out, from start to finish, in paragraph form.
    \marginnote{Write out your answer in paragraph form.  Use many different words to link sentences.  This is how you begin to get better at completing proofs.}
\end{enumerate}

